Giant Low Surface Brightness Galaxies As Merger Remnants

### 2018-05-04 glsb.tex
cool new sim citation

In \S\ref{sec:obs} we present new observations of Malin~2 and UGC~6614. In \S\ref{sec:compare_gals}, we compare the general properties of GLSB galaxies to sample of galaxies drawn from SDSS and the THINGS survey.  In \S\ref{sec:kinematics} and \S\ref{sec:sfh} we measure the stellar kinematics and star formation histories from our spectra. In \S\ref{sec:discussion}, we discuss what our observations imply for the possible formation mechanisms for GLSB systems.

@@

One of the most important features we can see in our data is the sharp metallicity gradient as one transitions between the central region and the disk in both Malin~2 and UGC~6614 (Figures~\ref{metal_malin2} and~\ref{metal_ugc}).  For both galaxies, the best fitting star formation histories show the stellar populations are dominated by old stars inside and outside the transition, but the metallicity changes by $\sim0.7-1.0$ dex.  This is a very sharp gradient compared to what has been found in normal spirals for the bulge-to-disk transition region.  \citet{MacArthur09} spectroscopically observed a range of spiral galaxies and found the metallicity profiles which were dropping, rising, and flat.  Using broadband colors, \citet{MacArthur04} also found metallicity gradients were generally very shallow, with steep metallicity gradients only in very late-type spiral galaxies. Also note that our metallicity gradient could be an underestimate of the true gradient since the central metallicities are at the limits of our model grid (i.e., they could have metallicities greater than [Z/H]=0.7). We have also spatially binned data, possibly washing out how sharp the transition is between bulge and disk.

\citet{Sanchez14} measure the Oxygen abundance from gas emission lines in a large sample of spiral galaxies and find a characteristic slope of $\alpha_{\rm{O/H}} = -0.1$\ dex $r_e^{-1}$ between 0.3 and 2 effective radii.  The metallicity gradients we observe in Malin~2 and UGC~6614 are factors of 2-4 steeper than the typical spiral. Also, of the 193 galaxies in \citet{Sanchez14} only four have a slope steeper than -0.3 \ dex $r_e^{-1}$ \citep[Figure~6]{Sanchez14}.  In their Figure~9, they combine the data to show a typical gas metallicity curve that drops linearly until $r_e\sim2$ and then flattens to a constant value.  This shape is very different from the step function change we observe in the GLSBs.

These are incredibly steep metallicity gradients.  The logarithmic gradients $\delta \rm{[Z/H]}\ \log{R/R_e}$ is $\sim -1.6$ for Malin~2 and $\sim -0.68$ for UGC~6614.  These are very steep compared to gradients typically observed in elliptical galaxies \citep{Greene12, Greene15}.  Other cites for elliptical gradients \citep{Kuntschner10}

\citet{Wilkinson15} find -0.15 dex/$R_e$ gradients for early type galaxies and flat metallicity gradients for late type galaxies.

@@

\citet{Kasparova14} present one of the most in depth observational studies of Malin~2.  Mass-modeling the system, they find that a huge isothermal core with $r_c=27.3$ kpc.  \citet{Kasparova14} also measure gas metallicities and find the metallicity drops from solar in the center of the system to half-solar at $r=30$\ kpc.  It is probably foolish to directly compare gas and stellar metallicities, but it is interesting that the gas metallicity gradient appears to be much lower than the stellar gradient.  This is probably consistent with a low metallicity gas-rich galaxy merging with an elliptical system that has much higher stellar metallicity, although a detailed simulation would be helpful in confirming what the final expected gas metallicity gradient should be.

@@

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
%\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Upper panels show kinematics from individual fibers, lower panels show the results after fibers have been binned into triples.  Left:  Stellar velocities measured with pPXF \citep{Cappellari04}. Middle:  Best fitting tilted ring model from kinemetry \citet{Krajn06}. Right: Best fit residuals.  \label{malin_kinem_fibers}}
\end{figure*}

\begin{figure*}
\epsscale{0.25}
%\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
%\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Rotation curves extracted from the best fitting models in Fig~\ref{malin_kinem_fibers}.  The dashed curves show the expected stellar mass contribution to the rotation curve.  Vertical lines mark $R_e$ and $5R_e$.  \HI\ data from \citet{Pickering97}.  \label{malin_kinem}}
\end{figure*}

@@

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataUGC06614.eps}\plotone{Plots/modelUGC06614.eps}\plotone{Plots/data_modelUGC06614.eps} \\
\plotone{Plots/databinned_UGC06614.eps}\plotone{Plots/modelbinned_UGC06614.eps}\plotone{Plots/data_modelbinned_UGC06614.eps}
\epsscale{1}
\caption{Upper panels show kinematics from individual fibers, lower panels show the results after fibers have been binned into triples.  Left:  Stellar velocities measured with pPXF \citep{Cappellari04}. Middle:  Best fitting tilted ring model from kinemetry \citet{Krajn06}. Right: Best fit residuals.  \label{ugc_kinem_fibers}}
\end{figure*}

\begin{figure*}
\epsscale{1}
\plottwo{Plots/kinembinned_UGC06614.eps}{Plots/kinem_HI_binned_UGC06614.eps}
\caption{Rotation curves extracted from the best fitting model in Fig~\ref{ugc_kinem_fibers}.  The dashed curve shows the expected stellar mass contribution to the rotation curve.  Vertical lines mark $R_e$ and $5R_e$.  \HI\ data from \citet{Pickering97}. \label{ugc_kinem}}
\end{figure*}

>>>>

In \S\ref{sec:obs} we present new observations of Malin~2 and UGC~6614. In \S\ref{sec:compare_gals}, we compare the general properties of GLSB galaxies to samples of galaxies drawn from the Sloan Digital Sky Survey (SDSS) and  The \HI Nearby Galaxy Survey (THINGS).  In \S\ref{sec:kinematics} and \S\ref{sec:sfh} we measure the stellar kinematics and star formation histories from our spectra. In \S\ref{sec:discussion}, we discuss what our observations imply for the possible formation mechanisms for GLSB systems.

@@

One of the most important features we can see in our data is the sharp metallicity gradient as one transitions between the central region and the disk in both Malin~2 and UGC~6614 (Figures~\ref{metal_malin2} and~\ref{metal_ugc}).  For both galaxies, the best fitting star formation histories show the stellar populations are dominated by old stars inside and outside the transition, but the metallicity changes by $\sim0.7-1.0$ dex.  This is a very sharp gradient compared to what has been found in normal spirals for the bulge-to-disk transition region.  \citet{MacArthur09} spectroscopically observed a range of spiral galaxies and found the metallicity profiles which were dropping, rising, and flat.  Using broadband colors, \citet{MacArthur04} also found metallicity gradients were generally very shallow, with steep metallicity gradients only in very late-type spiral galaxies. Also note that our metallicity gradient could be an underestimate of the true gradient since the central metallicities are at the limits of our model grid (i.e., they could have metallicities greater than [Z/H]=0.7). We have also spatially binned data, possibly diluting how sharp the transition is between bulge and disk.

\citet{Sanchez14} measure the Oxygen abundance from gas emission lines in a large sample of spiral galaxies and find a characteristic slope of $\alpha_{\rm{O/H}} = -0.1$\ dex $r_e^{-1}$ between 0.3 and 2 effective radii.  The metallicity gradients we observe in Malin~2 and UGC~6614 are factors of 2-4 steeper than the typical spiral. Also, of the 193 galaxies in \citet{Sanchez14} only four have a slope steeper than -0.3 \ dex $r_e^{-1}$ \citep[Figure~6]{Sanchez14}.  In their Figure~9, they combine the data to show a typical gas metallicity curve that drops linearly until $r_e\sim2$ and then flattens to a constant value.  This shape is very different from the step function change we observe in GLSBs.

These are incredibly steep metallicity gradients.  The logarithmic gradients $\delta \rm{[Z/H]}\ \log{R/R_e}$ is $\sim -1.6$ for Malin~2 and $\sim -0.68$ for UGC~6614.  These are very steep compared to gradients typically observed in elliptical galaxies \citep{Kuntschner10, Greene12, Greene15}.  \citet{Wilkinson15} find -0.15 dex/$R_e$ gradients for early type galaxies and flat metallicity gradients for late type galaxies.

XXX-While metallicity gradients this steep have not been observed, they do appear in simulations. \citet{Sanderson17} find that for their simulated Milky Way mass galaxies that a sharp break in stellar metallicity corresponds to the galactic radius where stellar populations become dominated by accretion. XXX--Should take a stab at saying how large the jump is.

@@

\citet{Kasparova14} present one of the most in depth observational studies of Malin~2.  Mass-modeling the system, they find that a huge isothermal core with $r_c=27.3$ kpc best fits the dynamics.  \citet{Kasparova14} also measure gas metallicities and find the metallicity drops from solar in the center of the system to half-solar at $r=30$\ kpc.  It is probably foolish to directly compare gas and stellar metallicities, but it is interesting that the gas metallicity gradient appears to be much lower than the stellar gradient.  This is probably consistent with a low metallicity gas-rich galaxy merging with an elliptical system that has much higher stellar metallicity, although a detailed simulation would be helpful in confirming what the final expected gas metallicity gradient should be.

@@

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
%\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Upper panels show kinematics from individual fibers, lower panels show the results after fibers have been binned into triples.  Left:  Stellar velocities measured with pPXF \citep{Cappellari04}. Middle:  Best fitting tilted ring model from kinemetry \citep{Krajn06}. Right: Best fit residuals.  \label{malin_kinem_fibers}}
\end{figure*}

\begin{figure*}
\epsscale{0.25}
%\plotone{Plots/dataMalin2.eps}\plotone{Plots/modelMalin2.eps}\plotone{Plots/data_modelMalin2.eps} \\
%\plotone{Plots/databinned_Malin2.eps}\plotone{Plots/modelbinned_Malin2.eps}\plotone{Plots/data_modelbinned_Malin2.eps}
%\epsscale{.5}
\epsscale{1}
\plottwo{Plots/kinembinned_Malin2.eps}{Plots/kinem_HI_binned_Malin2.eps}
\caption{Rotation curves extracted from the best fitting models in Fig~\ref{malin_kinem_fibers}.  The dashed curves show the expected stellar mass contribution to the rotation curve.  Vertical lines mark $R_e$ and $5R_e$.  Right panel includes \HI\ data from \citet{Pickering97}.  \label{malin_kinem}}
\end{figure*}

@@

\begin{figure*}
\epsscale{0.25}
\plotone{Plots/dataUGC06614.eps}\plotone{Plots/modelUGC06614.eps}\plotone{Plots/data_modelUGC06614.eps} \\
\plotone{Plots/databinned_UGC06614.eps}\plotone{Plots/modelbinned_UGC06614.eps}\plotone{Plots/data_modelbinned_UGC06614.eps}
\epsscale{1}
\caption{Upper panels show kinematics from individual fibers, lower panels show the results after fibers have been binned into triples.  Left:  Stellar velocities measured with pPXF \citep{Cappellari04}. Middle:  Best fitting tilted ring model from kinemetry \citep{Krajn06}. Right: Best fit residuals.  \label{ugc_kinem_fibers}}
\end{figure*}

\begin{figure*}
\epsscale{1}
\plottwo{Plots/kinembinned_UGC06614.eps}{Plots/kinem_HI_binned_UGC06614.eps}
\caption{Rotation curves extracted from the best fitting model in Fig~\ref{ugc_kinem_fibers}.  The dashed curve shows the expected stellar mass contribution to the rotation curve.  Vertical lines mark $R_e$ and $5R_e$.  Right panel includes \HI\ data from \citet{Pickering97}. \label{ugc_kinem}}
\end{figure*}
